---
title: "Example thermal-tolerance (TLS) analysis from Fv/Fm heat-stress data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = TRUE, fig.width = 8, fig.height = 6)

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(freqTLS)
library(kableExtra)

theme_set(theme_bw())
```

## Purpose

This notebook fits thermal load sensitivity (TLS) dose-response curves to heat-stress
Fv/Fm data using `freqTLS`, to complement the Fv/Fm extraction pipeline
(`FvFm_pipeline.py` / `FvFm_GUI.py`) in this repo. `FvFm_auto` (this repo's
automated Fo/Fm extraction) was used to generate the source values throughout.

It runs on `Data/tls_example_deidentified.csv` — a minimal, de-identified
dataset prepared from the fuller raw source data specifically so this example
can ship alongside the publication without the raw file's identifying
columns. See [Data provenance] for exactly how it was derived. Samples with
a pre-heat baseline Fv/Fm below 0.5 (not a "healthy" initial PSII state) were
already excluded during that preparation.

Species identity is de-identified throughout: `sp_id` (`SP1`–`SP5`) is
derived from `sp`, and `plant_id` is a simple per-species sequential code
(`SP1_1`, `SP1_2`, ...).

Threshold temperatures below follow the TDT convention: TX is
the temperature at which X% of the PSII function has been lost — so
**T90 is a harsher condition than T10** (90% vs. 10% decline). CTmax is a
related but distinct quantity: the curve's own *relative* midpoint,
`(low + up) / 2`, not a fixed absolute percentage — see
[Threshold temperatures (TX)].

The response variable itself — absolute Fv/Fm vs Fv/Fm expressed as a
proportion of each plant's own pre-heat baseline — is a modelling
choice with different implications for cross-species comparison; see
[Response variable: proportional vs. absolute Fv/Fm] for the reasoning
behind using the proportional response as the primary analysis here, with
absolute Fv/Fm retained as a comparison.

## Data provenance

This chunk is not re-run here (`eval=FALSE`) — it documents exactly how
`Data/tls_example_deidentified.csv` was produced from the raw source file.

```{r provenance, eval=FALSE}
raw <- read.csv("Data/subset_fvfm_doses.csv", stringsAsFactors = FALSE)

dat <- raw |>
  mutate(sp_id = paste0("SP", as.integer(factor(sp, levels = sort(unique(sp)))))) |>
  group_by(sp_id) |>
  mutate(plant_id = paste0(sp_id, "_", dense_rank(paste(chamber, ch_rep, Rep)))) |>
  ungroup() |>
  transmute(
    sp_id, plant_id,
    temp_C = Temp,
    time_min = Time,
    FvFm_initial = initial_fvfm,  # pre-heat baseline
    FvFm_post    = post_fvfm,     # 2h post-heat
    FvFm_final   = final_fvfm     # 24h / overnight
  ) |>
  filter(FvFm_initial >= 0.5)

write.csv(dat, "Data/tls_example_deidentified.csv", row.names = FALSE)
```

## Load and prepare data

```{r load-data}
dat <- read.csv("Data/tls_example_deidentified.csv", stringsAsFactors = FALSE)
glimpse(dat)

dur_breaks <- sort(unique(dat$time_min))
```

### Baseline QC

```{r qc-baseline}
dat |>
  group_by(sp_id) |>
  summarise(mean_i = mean(FvFm_initial), sd_i = sd(FvFm_initial), n = n(), .groups = "drop")
```

```{r qc-baseline-plot}
ggplot(dat, aes(temp_C, FvFm_initial, colour = sp_id)) +
  geom_jitter(width = 1, alpha = 0.4, size = 1) +
  scale_colour_viridis_d(option = "mako", end = 0.8) +
  labs(x = "Assigned temperature (°C)", y = "Baseline Fv/Fm", colour = "Species")
```

Baseline Fv/Fm is flat across assigned treatment, as expected — the pre-heat
baseline shouldn't depend treatment because it hasn't been applied yet.
Samples with baseline Fv/Fm below 0.5 were removed during data preparation 
(see [Data provenance]), so the minimum here is 0.5 by construction. Samples
vary within a healthy range (SD 0.02–0.06 per species, and species means 
differ by up to ~0.07). Likely this is a combination of biological variation, 
and leaf-to-leaf handling during sampling (e.g. how/where a disc is cut).

## Response variable: proportional vs. absolute Fv/Fm

Two ways to define the survival response for the same underlying data:

- **Absolute Fv/Fm** at 24h (or 2h) — used directly as the TLS response.
- **Proportional Fv/Fm** — that same reading divided by the plant's own
  pre-heat baseline (`FvFm_final / FvFm_initial`), i.e. the fraction of
  *that individual's own* capacity retained.

There are three considerations for these data:

1. **Sample-to-sample baseline variance.** The baseline QC above shows real
   spread in pre-heat Fv/Fm, both within and between species, plausibly
   including handling/cutting artefacts rather than biology. Absolute
   Fv/Fm at 24h carries that noise into the response; dividing by each 
   plant's own baseline removes it, at the cost of also propagating any
   noise in the baseline reading itself multiplicatively into the ratio.
   With species means differing by up to ~0.07 here, this could be large.
2. **`up` identifiability.** `freqTLS`'s own diagnostics (`check_tls()`)
   flag the upper asymptote as unreliable when the maximum observed
   survival in the data doesn't get close to 1 (in practice, needs to
   approach roughly 0.8+) — and that requirement gets harder to satisfy
   once `up` is allowed to vary *by species*, since each species-specific
   ceiling then needs its own near-undamaged observations. Absolute Fv/Fm
   tops out near ~0.8 by definition (close to a healthy leaf's maximum), 
   so `up` sits under the useful range for any per-species asymptote. 
   Dividing by each plant's own baseline rescales the *achievable* ceiling 
   to ~1 for every plant, which should make `up` (and any per-species version 
   of it) far better identified in the fitted models.
3. **Field of research familiarity.** Most people who work with chlorophyll
   fluorescence think in terms of raw Fv/Fm on its conventional ~0–0.85
   scale, not a 0–1 proportion of baseline term. Therefore, we retain absolute
   Fv/Fm on the raw data scale as a secondary analysis.

Given (1) and (2), the proportional response is used as the primary
analysis from here on; absolute Fv/Fm is refit with the same model
structure as a comparison once that structure is established through 
model selection.

```{r prep-primary}
o_dat <- dat |>
  mutate(prop_final = FvFm_final / FvFm_initial)

cat("prop_final range:", range(o_dat$prop_final), " | n == 0:", sum(o_dat$prop_final == 0),
    " | n > 1:", sum(o_dat$prop_final > 1), "\n")

o_dat <- o_dat |>
  mutate(prop_clamped = pmin(pmax(prop_final, 1e-4), 1 - 1e-4))
```

Unlike the absolute response (floored only, since it can't exceed ~0.85),
the proportional response needs clamping at **both** ends. It can hit exactly
0 (complete damage after 24 h) and can slightly exceed 1 (a leaf tissue sample
reading marginally higher at 24 h than its own pre-heat baseline — typically 
measurement noise, not real biology).

## Design & replication check

```{r replication-check}
dat |> distinct(sp_id, plant_id) |> count(sp_id, name = "n_individuals")

dat |>
  count(temp_C, time_min) |>
  pivot_wider(names_from = time_min, values_from = n) |>
  arrange(temp_C)
```

Each species has 6–7 individual plants, each measured across almost the
full 7 (temperature) x 5 (duration) grid — a median of 6 individuals per
species at any given treatment cell (the 30°C and 56°C edges, and the
120-minute column above 44°C, are intentionally reduced, since those
combinations are the least informative).

Each individual plant also contributes ~28 rows here (one row per treatment
thermal load it was measured at) — those are repeated measurements on the 
same physical plant, not independent samples, so a random intercept per
`plant_id` is still needed underneath the species fixed effect to avoid
pseudo-replication.

## Choosing a model structure

Following the standard workflow, model structure is decided by comparison,
not assumption. `t_ref = 60` minutes throughout; `method = "wald"` for fast
exploratory intervals. Switch to `"profile"`/`"bootstrap"` for a more robust
statistical model fit, but expect compute time to significantly increase.

A few candidate models are fitted to determine the appropriate structure, 
and these are evaluated using Akaike's Information Criterion (AIC).

```{r standardize}
dat_std <- standardize_data(
  o_dat, temp = "temp_C", duration = "time_min", proportion = "prop_clamped",
  duration_unit = "minutes", random_effects = c("sp_id", "plant_id")
)
```

```{r model-comparison}
fit_pop      <- fit_4pl(dat_std, family = "beta", t_ref = 60, method = "wald", quiet = TRUE)
fit_sp_re    <- fit_4pl(dat_std, ctmax = ~1 + (1|sp_id), z = ~1 + (1|sp_id),
                         family = "beta", t_ref = 60, method = "wald", quiet = TRUE)
fit_sp_fix   <- fit_4pl(dat_std, ctmax = ~0 + sp_id, z = ~0 + sp_id,
                         family = "beta", t_ref = 60, method = "wald", quiet = TRUE)
fit_sp_plant <- suppressWarnings(fit_4pl(dat_std, ctmax = ~0 + sp_id + (1|plant_id), z = ~0 + sp_id,
                         family = "beta", t_ref = 60, method = "wald", quiet = TRUE))
fit_sp_up    <- suppressWarnings(fit_4pl(dat_std, ctmax = ~0 + sp_id + (1|plant_id), z = ~0 + sp_id,
                         up = ~0 + sp_id, family = "beta", t_ref = 60, method = "wald", quiet = TRUE))

tibble::tibble(
  model = c("population only", "species random intercept", "species fixed effect",
            "+ plant random intercept (CTmax)", "+ up varying by species"),
  AIC   = c(AIC(fit_pop$fit), AIC(fit_sp_re$fit), AIC(fit_sp_fix$fit),
            AIC(fit_sp_plant$fit), AIC(fit_sp_up$fit))
) |> arrange(AIC)
```

Species matters a great deal, a **fixed** species effect is more appropriate 
than a random intercept given the replication here, and the `plant_id` random 
intercept matters too, given the repeated-measures structure. Unlike the 
absolute Fv/Fm version of this model (see [Comparison with absolute Fv/Fm]), 
a `plant_id` random intercept on `z` as well as `CTmax` does **not** converge 
here (`NA/NaN gradient`) — so `z` stays fixed-by-species only.

Note that this model structure protects `CTmax` from pseudo-replication but `z` 
has no plant_id term (the model did not converge), so its standard errors 
implicitly assume each plant's repeated measurements are independent for curve 
steepness. Any real plant-to-plant variation in `z` is absorbed into the 
residual dispersion (`phi`) rather than properly partitioned out, likely 
reducing `z`'s' reported uncertainty.

**Explore whether `k` or `low` should vary by species** 

```{r shape-by-species-test}
fit_sp_k   <- suppressWarnings(fit_4pl(dat_std, ctmax = ~0+sp_id+(1|plant_id), z = ~0+sp_id,
                     k = ~0+sp_id, family = "beta", t_ref = 60, method = "wald", quiet = TRUE))
fit_sp_low <- suppressWarnings(fit_4pl(dat_std, ctmax = ~0+sp_id+(1|plant_id), z = ~0+sp_id,
                     low = ~0+sp_id, family = "beta", t_ref = 60, method = "wald", quiet = TRUE))
fit_sp_upk <- suppressWarnings(fit_4pl(dat_std, ctmax = ~0+sp_id+(1|plant_id), z = ~0+sp_id,
                     up = ~0+sp_id, k = ~0+sp_id, family = "beta", t_ref = 60, method = "wald", quiet = TRUE))

tibble::tibble(
  model = c("up only (adopted)", "k only", "low only", "up + k together"),
  AIC   = c(AIC(fit_sp_up$fit), AIC(fit_sp_k$fit), AIC(fit_sp_low$fit), AIC(fit_sp_upk$fit))
) |> arrange(AIC)
```

`k` and `low` varying individually give no material AIC improvement over
`up` alone. `up` is the one shape parameter worth letting vary by species;
`k` and `low` stay shared. This is the direct CTmax/z parameterisation 
(not a reduced or reparameterised form), which — with this design (5 
species x 6–7 plants x a near-full temperature x duration grid) —
recovers real per-species differences in the quantities of scientific
interest without over-fitting the harder-to-identify shape parameters.

```{r fit-final}
fit_sp <- fit_sp_up
check_tls(fit_sp)
```

There are no diagnostic warnings: the proportional response resolves the
`up_not_approached` warning that the absolute-response model triggers 
(see [Comparison with absolute Fv/Fm]), simply by rescaling every 
plant's achievable ceiling to ~1 instead of ~0.85.

### Model summary

```{r summary-final}
summary(fit_sp)
```

Reading this table:

- **`low`** — shared floor survival (the residual signal in completely
  destroyed tissue), on the natural (0–1) scale.
- **`up:SPx`** — each species' own undamaged ceiling, expressed as a
  proportion of that species' own baseline.
- **`k`** — shared steepness of the decline (log10-time scale).
- **`CTmax:SPx`** — each species' relative-midpoint threshold temperature
  (°C) at `t_ref = 60` min — the headline dose-response quantity. Under the
  proportional response this is "the temperature at which a species has
  lost half the distance between its own floor and its own ceiling".
- **`z:SPx`** — each species' thermal sensitivity (°C per 10x change in
  exposure duration).
- **`phi`** — Beta-distribution dispersion (precision) parameter.
- **`sigma_CTmax`** — the estimated `plant_id` random-intercept standard
  deviation: how much individual plants vary around their species' `CTmax`
  (°C), after accounting for species and treatment.

```{r summary-table-reportable}
summary(fit_sp)$coefficients |>
  mutate(
    Parameter = sub(":.*$", "", parameter),
    Species   = ifelse(is.na(group), "—", group),
    Estimate  = sprintf("%.3f", estimate),
    SE        = sprintf("%.3f", std.error),
    `z value` = sprintf("%.2f", statistic),
    `p value` = ifelse(p.value < 0.001, "< 0.001", sprintf("%.3f", p.value))
  ) |>
  select(Parameter, Species, Estimate, SE, `z value`, `p value`) |>
  kbl(align = "llrrrr",
      caption = "Fitted 4PL parameters — primary model (proportional Fv/Fm response)") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) |>
  row_spec(0, bold = TRUE)
```

### Headline quantities, per species

```{r tls-species}
tls(fit_sp, by = "sp_id", method = "wald")$summary |>
  pivot_wider(names_from = quantity, values_from = c(median, lower, upper))
```

```{r get-shape}
get_shape(fit_sp)  # low/k/phi only — up is per-species now, see summary() above
```

### Threshold temperatures (TX)

```{r derive-thresholds}
thresholds <- map_dfr(paste0("SP", 1:5), function(s) {
  tibble::tibble(
    sp_id = s,
    T50 = suppressWarnings(derive_ctmax(fit_sp, surv = 0.5, group = s)),
    T90 = suppressWarnings(derive_ctmax(fit_sp, surv = 0.1, group = s))
  )
})
thresholds
```

### Plots

```{r plot-primary-curves}
plot_survival_curves(fit_sp) +
  scale_x_log10(breaks = dur_breaks) +
  labs(title = NULL, subtitle = NULL,
       x = "Exposure duration (min, log10 scale)",
       y = "Fv/Fm proportion of baseline (24h / initial)")
```

```{r plot-primary-surface}
plot_survival_surface(fit_sp, temps = seq(35, 50, 0.5)) +
  scale_y_log10(breaks = dur_breaks) +
  labs(title = NULL, subtitle = NULL,
       y = "Exposure duration (min, log10 scale)",
       fill = "Proportion\nof baseline\n(24h)")
```

### Time to T50, per species

```{r plot-tdt-curve-50}
suppressWarnings(plot_tdt_curve(fit_sp, p = 0.5, temps = seq(40, 52, length.out = 100))) +
  scale_y_log10(breaks = c(1, 5, 15, 30, 60, 120, 240, 480)) +
  labs(title = NULL, subtitle = NULL, colour = "Species",
       y = "Time to T50 (min, log10 scale)",
       caption = "Duration at which fitted Fv/Fm (24h, proportion of baseline) crosses 50% absolute survival, per species/temperature.")
```

Temperatures below 40°C are excluded for plotting these: at milder 
conditions the model's implied time-to-T50 extrapolates to durations
far beyond the tested window.

### Time to T90, per species

```{r plot-tdt-curve-90}
suppressWarnings(plot_tdt_curve(fit_sp, p = 0.1, temps = seq(40, 52, length.out = 100))) +
  scale_y_log10(breaks = c(1, 5, 15, 30, 60, 120, 240, 480)) +
  labs(title = NULL, subtitle = NULL, colour = "Species",
       y = "Time to T90 (min, log10 scale)",
       caption = "Duration at which fitted Fv/Fm (24h, proportion of baseline) crosses 90% absolute survival, per species/temperature.")
```

Temperatures below 40°C are excluded for plotting these: at milder 
conditions the model's implied time-to-T50 extrapolates to durations
far beyond the tested window.

## Comparison with absolute Fv/Fm

The same model structure (`ctmax = ~0+sp_id+(1|plant_id)`, `z = ~0+sp_id`,
`up = ~0+sp_id`), refit to absolute `FvFm_final` instead. This is a directly
comparable check of whether the response-variable choice changes conclusions.

```{r fit-absolute-comparison}
abs_dat <- dat |> mutate(FvFm_clamped = pmax(FvFm_final, 1e-4))
dat_std_abs <- standardize_data(abs_dat, temp = "temp_C", duration = "time_min", proportion = "FvFm_clamped",
                                 duration_unit = "minutes", random_effects = c("sp_id", "plant_id"))
fit_abs <- suppressWarnings(fit_4pl(dat_std_abs, ctmax = ~0+sp_id+(1|plant_id), z = ~0+sp_id, up = ~0+sp_id,
                    family = "beta", t_ref = 60, method = "wald", quiet = TRUE))
check_tls(fit_abs)
```

The absolute-response model **does** trigger `up_not_approached` warning.

```{r compare-abs-prop}
cmp <- bind_rows(
  tls(fit_sp,  by = "sp_id", method = "wald")$summary |> mutate(response = "Proportional (final/initial)"),
  tls(fit_abs, by = "sp_id", method = "wald")$summary |> mutate(response = "Absolute Fv/Fm (24h)")
) |> filter(quantity == "CTmax")

cmp |> select(sp_id, response, median, lower, upper) |>
  pivot_wider(names_from = response, values_from = c(median, lower, upper))

ggplot(cmp, aes(sp_id, median, colour = response)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 0.4)) +
  labs(x = "Species", y = "CTmax (°C)", colour = NULL) +
  theme(legend.position = "bottom")
```

`CTmax` estimates and species ranking are essentially unchanged from the
proportional to the absolute approach (largest shift ~0.4°C, well inside 
every CI; Spearman rank correlation = 1). The response variable choice 
here does not change which species is more or less heat tolerant — but it
does change the model fits and exact interpretation of the parameters.

## Acute post-heat (2h) curve

The same model selection logic applied to the 2 h measurements, as a 
secondary comparison against the 24 h measurements. Under the proportional 
response, the 2 h data support exactly the same structure as the 24 h data.
The absolute response version, needs a reduced structure for the 2 h data.

```{r fit-secondary}
p_dat <- dat |>
  mutate(prop_post = pmin(pmax(FvFm_post / FvFm_initial, 1e-4), 1 - 1e-4))
dat_std_p <- standardize_data(
  p_dat, temp = "temp_C", duration = "time_min", proportion = "prop_post",
  duration_unit = "minutes", random_effects = c("sp_id", "plant_id")
)
fit_p <- suppressWarnings(fit_4pl(dat_std_p, ctmax = ~0 + sp_id + (1|plant_id), z = ~0 + sp_id,
                  up = ~0 + sp_id, family = "beta", t_ref = 60, method = "wald", quiet = TRUE))
check_tls(fit_p)
```

```{r plot-secondary}
plot_survival_curves(fit_p) +
  scale_x_log10(breaks = dur_breaks) +
  labs(title = NULL, subtitle = NULL,
       x = "Exposure duration (min, log10 scale)",
       y = "Fv/Fm proportion of baseline (2h / initial)")
```

```{r compare-p-o}
bind_rows(
  tls(fit_sp, by = "sp_id", method = "wald")$summary |> mutate(period = "overnight (24h)"),
  tls(fit_p,  by = "sp_id", method = "wald")$summary |> mutate(period = "post-heat (2h)")
) |>
  filter(quantity == "CTmax") |>
  select(sp_id, period, median, lower, upper) |>
  pivot_wider(names_from = period, values_from = c(median, lower, upper))
```

Every species shows a higher CTmax overnight than at 2 h — highlighting
consistent overnight recovery. This comparison is descriptive
(two separately-fit curves); [Formal test of the recovery effect] below
gives a statistical test of the same pattern.

## Formal test of the recovery effect

Stacking the 2 h and 24 h readings into one long dataset with a `period`
factor lets `period` enter the model as an explicit term.

A model with `sp_id x period` interaction (mirroring the per-species
structure above) was attempted first, under both the proportional and
absolute responses — both failed to converge due to overparameterization.
The population-level model below — `period` as a fixed effect, `plant_id` 
as a random intercept, no species term — is the model that converges, and
directly answers the question: "is there an overall recovery effect?" 
while accounting for repeated measures on the same plant.

```{r joint-period-model}
joint_dat <- bind_rows(
  dat |> transmute(sp_id, plant_id, temp_C, time_min, period = "post",  prop = FvFm_post  / FvFm_initial),
  dat |> transmute(sp_id, plant_id, temp_C, time_min, period = "final", prop = FvFm_final / FvFm_initial)
) |>
  mutate(period = factor(period, levels = c("post", "final")),
         prop_clamped = pmin(pmax(prop, 1e-4), 1 - 1e-4))

dat_std_joint <- standardize_data(
  joint_dat, temp = "temp_C", duration = "time_min", proportion = "prop_clamped",
  duration_unit = "minutes", random_effects = c("sp_id", "plant_id")
)

fit_joint <- fit_4pl(dat_std_joint, ctmax = ~1 + period + (1|plant_id), z = ~1 + period,
                      family = "beta", t_ref = 60, method = "wald", quiet = TRUE)
diagnose_tdt_fit(fit_joint)
check_tls(fit_joint)
tls(fit_joint, method = "wald")
```

`CTmax` is significantly higher at 24 h than at 2 h (non-overlapping Wald
intervals) — a formal, population-level confirmation of the recovery effect
visible in the per-species comparison above.

## Recovery from acute stress: quantifying 2 h → 24 h change

This section deliberately uses **absolute** Fv/Fm differences, not
proportions — the question here ("how much Fv/Fm was regained overnight?")
is naturally additive, not relative. Here we decompose recovery into acute
damage, recoved, and irreparable from repeated measures in the same discs.
Acute damage (`initial − post`) splits exactly into what recovered overnight 
(`final − post`) and what remained lost at 24 h (`initial − final`).

```{r decompose-recovery}
rec <- dat |>
  mutate(
    acute_damage = FvFm_initial - FvFm_post,
    recovered    = FvFm_final - FvFm_post,
    irreparable  = FvFm_initial - FvFm_final
  )
```

```{r plot-recovery-scatter}
ggplot(rec, aes(temp_C, recovered, colour = sp_id)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_jitter(width = 1, alpha = 0.5, size = 1) +
  geom_smooth(se = TRUE, span = 1.2) +
  facet_wrap(~sp_id) +
  scale_colour_viridis_d(option = "mako", end = 0.8) +
  labs(x = "Temperature (°C)", y = "Δ Fv/Fm (24h − 2h)") +
  theme(legend.position = "none")
```

Recovery peaks in the mid-range (roughly 35–44°C) and collapses toward zero
at the most severe treatments (≥48°C). This indicates that repair capacity 
is limited, exhausted, or tissue is destroyed by 2 h at the harshest doses.

```{r plot-recovery-decompose}
rec |>
  group_by(temp_C, time_min) |>
  summarise(recovered = mean(recovered), irreparable = mean(irreparable), .groups = "drop") |>
  pivot_longer(c(recovered, irreparable), names_to = "component", values_to = "value") |>
  mutate(component = factor(component, levels = c("irreparable", "recovered"),
                             labels = c("Irreparable (still lost at 24h)",
                                        "Recovered overnight (repairable)"))) |>
  filter(value > 0) |> # removed negative change from baseline in repairable
  ggplot(aes(factor(time_min), value, fill = component)) +
  geom_col() +
  scale_fill_viridis_d(option = "plasma", end = 0.8, direction = -1) +
  facet_wrap(~temp_C, nrow = 1) +
  labs(x = "Exposure duration (min)", y = "Mean Fv/Fm change from baseline", fill = NULL) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom")
```

Averaged across species and individuals, the repairable (blue) component is
a real, non-trivial share of total damage at mild-to-moderate treatments
(30–44°C) but is essentially exhausted from 48°C upward.

Here we generate the figure reported in the main manscript, using combinations
of plots shown above, with additional annotations.

```{r plot-publication, fig.height = 6.5, fig.width = 9}

p1 <-
  plot_survival_curves(fit_sp) +
  scale_x_log10(breaks = dur_breaks) +
  scale_colour_viridis_d(name = "Temperature\n(°C)",
                      option = "C", end = 0.9) +
  facet_wrap(~group, ncol = 5) +
  labs(title = NULL, subtitle = NULL, caption = NULL,
       x = bquote("Exposure duration (min," ~ log[10] ~ scale * ")"),
       y = bquote(atop(italic(F)[v] / italic(F)[m] ~ "proportion of", baseline ~ (24 ~ h / initial)))) +
  theme_bw()

p2 <-
  plot_survival_surface(fit_sp, temps = seq(35, 50, 0.5)) +
  scale_y_log10(breaks = dur_breaks) +
  scale_fill_viridis_c(name = bquote(atop(italic(F)[v] / italic(F)[m] ~ "proportion of", baseline ~ (24 ~ h / initial))),
                       option = "D",
                       limits = c(0, 1)) +
  facet_wrap(~group, ncol = 5) +
  labs(title = NULL, subtitle = NULL, caption = NULL,
       y = bquote(atop("Exposure duration", (min*"," ~ log[10] ~ scale)))) +
  theme_bw()

tls_labels <- tls(fit_sp, by = "sp_id", method = "wald")$summary |>
  pivot_wider(names_from = quantity, values_from = c(median, lower, upper)) |>
  transmute(group = sp_id,
            label = sprintf("CTmax = %.1f°C\nz = %.2f", median_CTmax, median_z))

p3 <-
  suppressWarnings(plot_tdt_curve(fit_sp, p = 0.5, temps = seq(40, 52, length.out = 100))) +
  scale_y_log10(breaks = c(1, 5, 15, 30, 60, 120, 240), limits = c(1,300)) +
  scale_x_continuous(breaks = c(40,44,48), limits = c(40,50)) +
  scale_colour_viridis_d(option = "mako", end = 0.8, name = "Species") +
  facet_wrap(~group, ncol = 5) +
  geom_text(data = tls_labels, aes(label = label), x = Inf, y = Inf,
            hjust = 1.05, vjust = 1.4, inherit.aes = FALSE, size = 2.8) +
  labs(title = NULL, subtitle = NULL, caption = NULL,
       y = bquote(atop("Time to"~italic(T)[50], (min*"," ~ log[10] ~ scale)))) +
  theme_bw()

ggpubr::ggarrange(p1, p2, p3, ncol = 1, nrow = 3, align = "hv", labels = "auto")
```